## Libraries

```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(data.table)

# spatial
library(sf)
library(rgdal)
library(rmapshaper)

# plotting
library(eulerr)
library(cowplot)
library(scales)
library(gridExtra)
library(grid)

setwd("C:/Users/Lizzy/Documents/GitHub/wa_inat_plants/")
```

## Read in Yosemite data

```{r}
#setwd("C:/Users/Lizzy/Documents/GitHub/wa_inat_plants/")

Yos_species_list <- read_csv("data_Yosemite/Yosemite_all_data_2024.08.28.csv")

Yos_species_paper_2 <- Yos_species_list %>%
  filter(voucher_END == "1" | iNat_RG_END == "1")

Yos_iNat <- read_csv("data_Yosemite/Yosemite_ResearchGrade.csv") %>%
  mutate(accepted_name = stringr::word(scientific_name, start = 1, end = 2))
  
Yosemite_synonym_lookup <- read_csv("data_Yosemite/Yosemite_synonyms_lookup_2024.06.11.csv") %>%
  select(scientificName = starting_name, current_name, name_source)

CCH2_data <- read_csv("data_Yosemite/Yosemite_CCH_records_2024.06.14.csv") %>%
  filter(phylum == "Tracheophyta") %>%                  # retain only vascular plants
  filter(keep_omit == c("keep")) %>%                    # retain only valid vouchers
  filter(year > 1859) %>%                               # remove garbage years
  filter(recordedBy != "Gill, Raymond J.") %>%          # remove collection with photographic observations only
  filter(stringr::str_detect(scientificName, " ")) %>%  # remove genera
  left_join(Yosemite_synonym_lookup) %>%
  select(current_name, scientificName, keep_omit, institutionCode, recordedBy_simple, eventDate, year, locality, institutionCode, 
         recordNumber, decimalLatitude, decimalLongitude, recordedBy, county) %>%
  filter(!is.na(current_name))

gbif_to_include <- read_csv("data_Yosemite/Yosemite_GBIF_records_2024.06.14.csv") %>%
  filter(keep_omit == "keep") %>% 
  rename(scientificName = scientific_name)

voucher_data <- CCH2_data %>%
  bind_rows(gbif_to_include)
```

## Read in Royal data

```{r}
Royal_observers <- read_csv("data_Royal/species_per_collector.csv")
```

## Figure 1. Maps of long unobserved, recent finds

```{r}
establish_colours <-  c("native" = "#6A51A3", "introduced" = "red")
source_shapes <-  c("iNat" = 17, "voucher" = 16)

Yos_old_only <- Yos_species_paper_2 %>%
  filter(is.na(recent_voucher) & is.na(iNat_RG_END)) %>%
  mutate(establishment_means = ifelse(establishment_means == "invasive", "introduced", establishment_means))

#old_voucher_locations <- CCH2_data %>%
#  filter(current_name %in% Yos_old_only$accepted_name, keep_omit == "keep") %>%
  # turning off so that I retain actual final location and will manually fill in locations
  #filter(!is.na(decimalLatitude)) %>%   
#  arrange(-year) %>%
#  group_by(current_name) %>%
#  summarise_all(.funs = first) %>%
#  distinct() %>%
#  ungroup() %>%
#  left_join(Yos_old_only %>% select(current_name = accepted_name, establishment_means))

#old_voucher_locations %>% write_csv("data_Yosemite/Yosemite_long_unseen_last_location.csv")

old_voucher_locations <- read_csv("data_Yosemite/long_unseen_edited_lat_lon.csv") %>%
  mutate(source = "voucher") %>%
  left_join(Yos_old_only %>% select(current_name = accepted_name, establishment_means))

Yos_recent_only <- Yos_species_paper_2 %>%
  filter(is.na(old_voucher) & (iNat_RG_END == "1" | recent_voucher == "1"))

iNat_recent_additions <- Yos_iNat %>%
  select(scientific_name, observed_on, latitude, longitude) %>%
  filter(scientific_name %in% Yos_recent_only$accepted_name) %>%
  mutate(source = "iNat")
  
CCH_recent_additions <- CCH2_data %>%
  filter(current_name %in% Yos_recent_only$accepted_name, keep_omit == "keep") %>%
  # turning off so that I retain actual final location and will manually fill in locations
  #filter(!is.na(decimalLatitude), decimalLatitude > 37) %>%
  arrange(year) %>%
  group_by(current_name) %>%
  summarise_all(.funs = first) %>%
  distinct() %>%
  ungroup() %>%
  select(scientific_name = current_name, observed_on = year, latitude = decimalLatitude , longitude = decimalLongitude, recordedBy_simple, locality, institutionCode) %>%
  mutate(observed_on = lubridate::make_date(observed_on, 1, 1))

#CCH_recent_additions %>% write_csv("data_Yosemite/Yosemite_CCH_recent_additions.csv")

CCH_recent_additions <- read_csv("data_Yosemite/CCH_recent_additions_latlon_fixed.csv") %>% 
  mutate(
    source = "voucher", 
    observed_on = date(observed_on)
    )

recent_additions_all <- iNat_recent_additions %>%
  bind_rows(CCH_recent_additions) %>%
  left_join(Yos_recent_only %>% select(scientific_name = accepted_name, establishment_means)) %>%
  arrange(observed_on) %>%
  group_by(scientific_name) %>%
  summarise_all(.funs = first) %>%
  ungroup() %>%
  distinct()

yosemite_outline <- st_read("data_Yosemite/Yosemite.shp") %>%
  st_transform(crs = 4326) %>%
  st_make_valid()

yos_roads <- st_read("data_Yosemite/Yosemite_roads.gpx", layer = "tracks") %>%
  st_transform(crs = 4326) %>%
  st_make_valid()

yos_roads_clipped <- rmapshaper::ms_clip(target = yos_roads, clip = yosemite_outline[1])

Yos_long_unseen <- 
  ggplot() + 
  geom_sf(mapping = aes(), data = yosemite_outline , stat = "sf", fill = "#C5DDC2", color = "grey50") +
  geom_sf(mapping = aes(), data = yos_roads_clipped, stat = "sf") +
  geom_point(aes(longitude, latitude, col = factor(establishment_means)), data = old_voucher_locations, position = position_jitter(height=.005, width=.005)) +
  scale_colour_manual(values = establish_colours, name = "origin") +
  #scale_x_continuous(expand = c(0, 0)) +
  #scale_y_continuous(expand = c(0, 0)) +
  theme_classic() + 
  annotation_scale(location = "br", width_hint = 0.5, pad_y = unit(0, "cm"), pad_x = unit(0, "cm"),line_width = 0.5, height = unit(0.2, "cm")) +
  ggtitle("b. Yosemite") +
  theme(
    legend.position = "none",
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank(),
    axis.ticks.length = unit(0, "pt"),
    plot.title = element_text(hjust = 0, size = 10),
    plot.margin=unit(c(.0,.0,.0,.0),"cm")
  )

Yos_new_discoveries_tmp <-
  ggplot() + 
  geom_sf(mapping = aes(), data = yosemite_outline , stat = "sf", fill = "#C5DDC2", color = "grey50") +
  geom_sf(mapping = aes(), data = yos_roads_clipped, stat = "sf") +
  geom_point(aes(longitude, latitude, col = factor(establishment_means), shape = factor(source)), data = recent_additions_all, position = position_jitter(height=.005, width=.005)) +
  scale_colour_manual(values = establish_colours, name = "species origin") +
  scale_shape_manual(values = source_shapes, name = "data source") +
  #scale_x_continuous(expand = c(0, 0)) +
  #scale_y_continuous(expand = c(0, 0)) +
  theme_classic() +
  ggtitle("d. Yosemite") +
  theme(
    legend.position = "bottom",
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank(),
    axis.ticks.length = unit(0, "pt"),
    plot.title = element_text(hjust = 0, size = 10),
    plot.margin=unit(c(.0,.0,.0,.0),"cm")
  )

legend_maps <- get_legend(Yos_new_discoveries_tmp) #+ theme(legend.margin = margin(2, 2, 2, 2))

Yos_new_discoveries <- Yos_new_discoveries_tmp + theme(legend.position = "none")
```

```{r}
royal_location <- read_csv("data_Royal/Royal_locations_first_late.csv")
establish_colours2 <-  c("native" = "#6A51A3", "introduced" = "red")
source_shapes <-  c("iNat" = 17, "voucher" = 16)

Royal_long_unseen_locations <- royal_location %>%
  filter(category == "long unseen", `record type` == "voucher")

Royal_recent_additions_all <- royal_location %>%
  filter(category == "new discovery") %>%
  rename(source = `record type`)

royal_outline <- st_read("data_Royal/Royal_outline.shp") %>%
  st_transform(crs = 4326) %>%
  st_make_valid()

royal_roads <- st_read("data_Royal/Royal_roads.gpx", layer = "tracks") %>%
  st_transform(crs = 4326) %>%
  st_make_valid()

royal_roads_clipped <- rmapshaper::ms_clip(target = royal_roads, clip = royal_outline[1])


Royal_long_unseen <-
  ggplot() + 
  geom_sf(mapping = aes(), data = royal_outline , stat = "sf", fill = "#C5DDC2", color = "grey50") +
  geom_sf(mapping = aes(), data = royal_roads_clipped, stat = "sf") +
  geom_point(aes(longitude, latitude, col = factor(establishment_means)), data = Royal_long_unseen_locations, position = position_jitter(height=.001, width=.001)) +
  scale_colour_manual(values = establish_colours2, name = "origin") +
  #scale_x_continuous(expand = c(0, 0)) +
  #scale_y_continuous(expand = c(0, 0)) +
  theme_classic() + 
  ggtitle("a. Royal") +
  theme(
    legend.position = "none",
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank(),
    axis.ticks.length = unit(0, "pt"),
    plot.title = element_text(hjust = 0, size = 10),
    plot.margin=unit(c(.0,.0,.0,.0),"cm")
  )

Royal_new_discoveries <-
  ggplot() + 
  geom_sf(mapping = aes(), data = royal_outline , stat = "sf", fill = "#C5DDC2", color = "grey50") +
  geom_sf(mapping = aes(), data = royal_roads_clipped, stat = "sf") +
  geom_point(aes(longitude, latitude, col = factor(establishment_means), shape = factor(source)), data = Royal_recent_additions_all, position = position_jitter(height=.001, width=.001)) +
  scale_colour_manual(values = establish_colours2, name = "species origin") +
  scale_shape_manual(values = source_shapes, name = "data source") +
  annotation_scale(location = "br", width_hint = 0.5, pad_y = unit(1.4, "cm"), pad_x = unit(.3, "cm"), line_width = 0.5, height = unit(0.2, "cm")) +
  #scale_x_continuous(expand = c(0, 0)) +
  #scale_y_continuous(expand = c(0, 0)) +
  theme_classic() +
  ggtitle("c. Royal") +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank(),
    axis.ticks.length = unit(0, "pt"),
    plot.title = element_text(hjust = 0, size = 10),
    plot.margin=unit(c(0,0,0,0),"cm"),
    legend.position = "none"
    )
```

```{r}
#Yos_maps_combined <- ggdraw(plot_grid(Yos_long_unseen, Yos_new_arrivals, ncol=1, align='v', rel_widths=c(1, 1)))

#blank <- grid::rectGrob(gp = gpar(fill="white", col = "white"))

unseen <- grid.arrange(
  Royal_long_unseen, Yos_long_unseen,
  ncol = 2,
  top = textGrob("Last collection location for long-unseen species", gp = gpar(fontsize = 10), hjust = 1.2)
  )

new_discoveries <- grid.arrange(
  Royal_new_discoveries, Yos_new_discoveries,
  ncol = 2,
  top = textGrob("First collection location for post-1990 discoveries", gp = gpar(fontsize = 10), hjust = 1.2)
  )

legend <- grid.arrange(
  legend_maps, blank,
  ncol = 2
  )

all_maps <- grid.arrange(
  unseen, new_discoveries, legend_maps,
  ncol = 1, 
  heights = c(1, 1, 0.08)
)

cowplot::save_plot("outputs/maps.jpg", all_maps, base_width = 8, base_height = 9)
```



## Figure 2. Decline in "unobserved" over time

### - generate data summaries - Yosemite

```{r}

paper2_fig2 <-
  Yos_species_paper_2 %>%
  select(accepted_name, establishment_means, voucher_END, iNat_RG_END, old_voucher, recent_voucher) %>%
  distinct()

# vouchers
recent_obs_CCH <- CCH2_data %>%
  filter(keep_omit == "keep", year > 1989) %>%
  filter(current_name %in% paper2_fig2$accepted_name) %>%
  select(accepted_name = current_name, year) %>%
  mutate(dataset = "CCH") %>%
  arrange(year) %>%
  distinct(accepted_name, .keep_all = TRUE)

summary_recent_CCH <- recent_obs_CCH %>%
  group_by(year) %>%
  mutate(re_finds_per_year = n()) %>%
  distinct(re_finds_per_year) %>%
  ungroup()

#write_csv(summary_recent_CCH, "data_Yosemite/summary_recent_CCH.csv")

# iNaturalist
recent_obs_iNat <- Yos_iNat %>%
  select(scientific_name, observed_on) %>%
  mutate(year = year(observed_on)) %>%
  filter(year > 2007) %>%
  filter(scientific_name %in% Yos_species_paper_2$accepted_name) %>%
  select(accepted_name = scientific_name, year) %>%
  mutate(dataset = "iNat") %>%
  arrange(year) %>%
  distinct(accepted_name, .keep_all = TRUE)
# have confirmed there were no species with pre-2008 observations that haven't been seen since - so filtering changes when species were "ticked off" the iNat list, but not whether they are ticked off the list, at least for Yosemite

summary_recent_iNat <- recent_obs_iNat %>%
  group_by(year) %>%
  mutate(re_finds_per_year = n()) %>%
  distinct(re_finds_per_year) %>%
  ungroup() 

#write_csv(summary_recent_iNat, "data_Yosemite/summary_recent_iNat.csv")

# combined voucher & iNat

recent_obs_combined <- recent_obs_CCH %>%
  bind_rows(recent_obs_iNat) %>%
  arrange(accepted_name, year) %>%
  mutate(accepted_name = stringr::word(accepted_name, start = 1, end = 2)) %>%
  group_by(accepted_name) %>%
  summarise_all(.funs = first) %>%
  ungroup() %>%
  arrange(year)

summary_recent_combined <- recent_obs_combined %>%
  group_by(year) %>%
  mutate(re_finds_per_year = n()) %>%
  distinct(re_finds_per_year) %>%
  ungroup()

#write_csv(summary_recent_combined, "data_Yosemite/summary_recent_combined.csv")

summary_2_lines <- recent_obs_combined %>%
  mutate(counts = 1) %>%
  pivot_wider(names_from = dataset, values_from = counts) %>%
  mutate(
    iNat = if_else(is.na(iNat), 0, iNat),
    CCH = if_else(is.na(CCH), 0, CCH)
    ) %>%
  select(-accepted_name) %>%
  group_by(year) %>%
  mutate(
    CCH_finds_per_year = sum(CCH),
    iNat_finds_per_year = sum(iNat)
    ) %>%
  select(-CCH, -iNat) %>%
  distinct() %>%
  ungroup()

#write_csv(summary_2_lines, "data_Yosemite/summary_2_lines.csv")
```

### - generate data summaries - Royal

```{r}
Royal_need_finding <- read_csv("data_Royal/need_to_be_found.csv")

Royal_recent_vouchers <- Royal_need_finding %>%
  filter(!is.na(First_voucher_since_1990)) %>%
  arrange(First_voucher_since_1990)

Royal_recent_vouchers_summary <- Royal_recent_vouchers %>%
  group_by(First_voucher_since_1990) %>%
  mutate(re_finds_per_year = n()) %>%
  distinct(re_finds_per_year) %>%
  ungroup()

#write_csv(Royal_recent_vouchers_summary, "data_Royal/Royal_recent_vouchers_summary.csv")

Royal_recent_iNat <- Royal_need_finding %>%
  filter(!is.na(First_iNat_since_1990)) %>%
  arrange(First_iNat_since_1990)

Royal_recent_iNat_summary <- Royal_recent_iNat %>%
  group_by(First_iNat_since_1990) %>%
  mutate(re_finds_per_year = n()) %>%
  distinct(re_finds_per_year) %>%
  ungroup()

#write_csv(Royal_recent_iNat_summary, "data_Royal/Royal_recent_iNat_summary.csv")

Royal_recent_combined <- Royal_need_finding %>%
  filter(!is.na(First_voucher_since_1990) | !is.na(First_iNat_since_1990)) %>%
  select(scientific_name, First_voucher_since_1990, First_iNat_since_1990) %>%
  pivot_longer(cols = 2:3) %>%
  filter(!is.na(value)) %>% 
  mutate(name = ifelse(stringr::str_detect(name,"iNat"), "iNat", "voucher")) %>%
  arrange(value) %>%
  distinct(scientific_name, .keep_all = TRUE) %>%
  rename(year = value, source = name) %>%
  mutate(counter = 1) %>%
  select(-scientific_name) %>%
  group_by(year, source) %>%
  mutate(
    counter = sum(counter)
    ) %>%
  distinct() %>%
  ungroup() %>%
  pivot_wider(names_from = source, values_from = counter)

#write_csv(Royal_recent_combined, "data_Royal/Royal_recent_combined.csv", na = "")
```


### - generate plots

```{r}
# try https://stackoverflow.com/questions/61775003/ggplot2-create-shaded-area-with-gradient-below-curve to get colors to grade; won't let you use gradients with geom_ribbon

summary_recent_CCH <- read_csv("data_Yosemite/summary_recent_CCH.csv")
summary_recent_iNat <- read_csv("data_Yosemite/summary_recent_iNat.csv")
summary_recent_combined <- read_csv("data_Yosemite/summary_recent_combined.csv")
summary_recent_2lines <- read_csv("data_Yosemite/summary_2_lines.csv")

Royal_recent_vouchers_summary <- read_csv("data_Royal/Royal_recent_vouchers_summary.csv") %>% rename(year = First_voucher_since_1990)
Royal_recent_iNat_summary <- read_csv("data_Royal/Royal_recent_iNat_summary.csv") %>% rename(year = First_iNat_since_1990)
Royal_combined <- read_csv("data_Royal/Royal_recent_combined.csv")

combined_plot_a <- ggplot() + 
  geom_rect(aes(xmin = 1990, xmax = 2025, ymin = 0, ymax = min(Royal_recent_vouchers_summary$remaining)), fill = "#608CB8", colour = NA, alpha = 0.1) + 
  geom_rect(aes(xmin = 2003, xmax = 2025, ymin = 0, ymax = min(Royal_recent_iNat_summary$remaining)), fill = "#74AC00", colour = NA, alpha = 0.1) + 
  geom_ribbon(aes(x = year, ymin = min(Royal_recent_vouchers_summary$remaining), ymax = remaining), data = Royal_recent_vouchers_summary, fill = "#608CB8", color = "#608CB8", alpha = 0.35) +
  geom_ribbon(aes(x = year, ymin = min(Royal_recent_iNat_summary$remaining), ymax = remaining), data = Royal_recent_iNat_summary, fill = "#74AC00", color = "#74AC00", alpha = 0.35) +
  scale_x_continuous(limits = c(1990, 2025), breaks = c(1990, 2000, 2010, 2020), expand = c(0,0)) +
  scale_y_continuous(limits = c(0, 1200), breaks = c(200, 400, 600, 800, 1000, 1200), expand = c(0,0)) +
  geom_text(aes(x = 1996, y = 435, label = "status", fontface = "italic")) +
  geom_text(aes(x = 1996, y = 365, label = "unknown", fontface = "italic")) +
  geom_text(aes(x = 2014, y = 300, label = "status unknown", fontface = "italic")) +
  geom_text(aes(x = 1997, y = 860, label = "vouchers only")) +
  geom_text(aes(x = 2012, y = 960, label = "iNaturalist only")) +
  theme_classic()  +
  labs(
    title = "Royal"
  ) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )

combined_plot_b <- ggplot() + 
    geom_rect(aes(xmin = 1990, xmax = 2025, ymin = 0, ymax = min(Royal_combined$combined_remaining)), fill = "#6EA387", colour = NA, alpha = 0.15) +
    geom_ribbon(aes(x = year, ymin = min(Royal_combined$combined_remaining), ymax = combined_remaining), data = Royal_combined %>% filter(year < 2020), fill = "#608CB8", alpha = 0.2) +
    geom_ribbon(aes(x = year, ymin = min(Royal_combined$combined_remaining), ymax = combined_remaining), data = Royal_combined %>% filter(year > 2002), fill = "#74AC00", alpha = 0.2) +
  geom_line(aes(x = year, y = combined_remaining), data = Royal_combined, color = "#74AC00", size = 1) +
  geom_line(aes(x = year, y = voucher_remaining), data = Royal_combined, color = "#608CB8", size = 1) +
  scale_x_continuous(limits = c(1990, 2025), breaks = c(1990, 2000, 2010, 2020), expand = c(0,0)) +
  scale_y_continuous(limits = c(0, 1200), breaks = c(200, 400, 600, 800, 1000, 1200), expand = c(0,0)) +
  geom_text(aes(x = 1996, y = 800, label = "combined")) +
  geom_text(aes(x = 1996, y = 730, label = "resources")) +
  geom_text(aes(x = 2008, y = 300, label = "status unknown", fontface = "italic")) +
theme_classic()  +
  labs(
    title = " "
  ) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )

combined_plot_c <- ggplot() + 
  geom_rect(aes(xmin = 1990, xmax = 2024, ymin = 0, ymax = min(summary_recent_CCH$remaining)), fill = "#608CB8", colour = NA, alpha = 0.1) + 
  geom_rect(aes(xmin = 2008, xmax = 2024, ymin = 0, ymax = min(summary_recent_iNat$remaining)), fill = "#74AC00", colour = NA, alpha = 0.1) + 
  geom_ribbon(aes(x = year, ymin = min(summary_recent_CCH$remaining), ymax = remaining), data = summary_recent_CCH, fill = "#608CB8", color = "#608CB8", alpha = 0.35) +
  geom_ribbon(aes(x = year, ymin = min(summary_recent_iNat$remaining), ymax = remaining), data = summary_recent_iNat, fill = "#74AC00", color = "#74AC00", alpha = 0.35) +
  scale_x_continuous(limits = c(1990, 2025), breaks = c(1990, 2000, 2010, 2020), expand = c(0,0)) +
  scale_y_continuous(limits = c(0, 1600), breaks = c(500, 1000, 1500), expand = c(0,0)) +
  geom_text(aes(x = 1999, y = 240, label = "status unknown", fontface = "italic")) +
  geom_text(aes(x = 2016, y = 370, label = "status unknown", fontface = "italic")) +
  geom_text(aes(x = 1998, y = 850, label = "vouchers only")) +
  geom_text(aes(x = 2015, y = 850, label = "iNaturalist only")) +
  #scale_y_log10(limits = c(1, 1800), breaks = c(1, 10, 100, 1000, 1577), expand = c(0,0)) +
  theme_classic()  +
  labs(
    title = "Yosemite"
  ) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )

combined_plot_d <- ggplot() + 
    geom_rect(aes(xmin = 1990, xmax = 2024, ymin = 0, ymax = min(summary_recent_2lines$combined_remaining)), fill = "#6EA387", colour = NA, alpha = 0.15) +
    geom_ribbon(aes(x = year, ymin = min(summary_recent_2lines$combined_remaining), ymax = combined_remaining), data = summary_recent_2lines %>% filter(year < 2013), fill = "#608CB8", alpha = 0.2) +
    geom_ribbon(aes(x = year, ymin = min(summary_recent_2lines$combined_remaining), ymax = combined_remaining), data = summary_recent_2lines %>% filter(year > 2007), fill = "#74AC00", alpha = 0.2) +
  #scale_fill_gradient(low = "#608CB8", high = "#74AC00") +
  geom_line(aes(x = year, y = combined_remaining), data = summary_recent_2lines, color = "#74AC00", size = 1) +
  geom_line(aes(x = year, y = CCH_remaining), data = summary_recent_2lines, color = "#608CB8", size = 1) +
  scale_x_continuous(limits = c(1990, 2025), breaks = c(1990, 2000, 2010, 2020), expand = c(0,0)) +
  scale_y_continuous(limits = c(0, 1600), breaks = c(500, 1000, 1500), expand = c(0,0)) +
  geom_text(aes(x = 1998, y = 850, label = "combined")) +
  geom_text(aes(x = 1998, y = 750, label = "resources")) +
  geom_text(aes(x = 2008, y = 110, label = "status unknown", fontface = "italic")) +
theme_classic()  +
  labs(
    title = " "
  ) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )



decline_in_unobserved <- grid.arrange(
  combined_plot_a, combined_plot_b, combined_plot_c, combined_plot_d,
  ncol = 2,
  left = textGrob("species without recorded observations since 1990", rot = 90, gp = gpar(fontsize = 10)),
  bottom = textGrob("year", gp = gpar(fontsize = 10))
)

cowplot::save_plot("outputs/decline_in_unobserved.jpg", decline_in_unobserved, base_width = 8, base_height = 8)

```

## Figure 3/4/S. "Finds per observer" histogram

### - generate Yosemite iNat summary

```{r}
iNat_discoveries <- Yos_species_list %>%
  filter(iNat_RG_END == "1", is.na(voucher_END))

iNat_discoverers <- Yos_iNat %>%
  select(scientific_name, observed_on, user_login) %>%
  arrange(observed_on) %>%
  mutate(scientific_name = stringr::word(scientific_name, start = 1, end = 2)) %>%
  filter(scientific_name %in% iNat_discoveries$accepted_name) %>%
  group_by(scientific_name) %>%
  summarise_all(.funs = first) %>%
  ungroup() %>%
  group_by(user_login) %>%
  mutate(counts_per_obs = n()) %>%
  ungroup() %>% 
  distinct(user_login, counts_per_obs) %>%
  mutate(
    source = "iNat",
    decade_plus = "1990 - 2024"
    ) %>%
  rename(first_observer = user_login)
```

### - generate Yosemite voucher summary

```{r}
counts_per_decade <- Yos_species_list %>%
  select(accepted_name, establishment_means, voucher_END, year_first, first_observer) %>%
  filter(voucher_END == "1") %>%
  arrange(year_first) %>%
  mutate(decade_plus = 
           case_when(
             year_first < 1900 ~ "1860-1900",
             year_first >= 1900 & year_first < 1930 ~ "1900 - 1930",
             year_first >= 1930 & year_first < 1990 ~ "1930 - 1990",
             year_first >= 1990  ~ "1990 - 2024"
             ),
         source = "voucher") %>%
  group_by(decade_plus, first_observer, source) %>%
  summarise(counts_per_obs = n()) %>%
  distinct(decade_plus, first_observer, counts_per_obs, source) %>%
  ungroup() %>%
  filter(!is.na(decade_plus)) %>%
  bind_rows(iNat_discoverers)
```

### - generate Royal summary

```{r}
Royal_observer_summary <- Royal_observers %>%
  rename(first_observer = first_collector, year_first = year_first_collection) %>%
  arrange(year_first) %>%
  mutate(decade_plus = 
           case_when(
             year_first < 1900 ~ "1860-1900",
             year_first >= 1900 & year_first < 1930 ~ "1900 - 1930",
             year_first >= 1930 & year_first < 1990 ~ "1930 - 1990",
             year_first >= 1990  ~ "1990 - 2024"
             )
         ) %>%
  group_by(decade_plus, first_observer, source) %>%
  summarise(counts_per_obs = n()) %>%
  distinct(decade_plus, first_observer, counts_per_obs, source) %>%
  ungroup() %>%
  filter(!is.na(decade_plus))
```

### - histograms of "finds" by observer

```{r}
collection_colours <-  c("iNaturalist" = "#74AC00", "voucher" = "#608CB8")

counts_per_decade <- counts_per_decade %>%
  mutate(source = ifelse(source == "iNat", "iNaturalist", source))

Yos_histograms <- ggplot(data = counts_per_decade, aes(x = counts_per_obs)) +
  geom_histogram(aes_string(fill = "source")) +
  facet_wrap(~ decade_plus, ncol = 1) +
  scale_x_log10() +
  scale_y_continuous(limits = c(0,40)) +
  scale_fill_manual(values = collection_colours, name = "data source") +
  labs(
    y = "number of observers",
    #x = "number of first records",
    title = "Yosemite"
    ) +
  theme(
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank()
    ) 

Royal_histograms <- ggplot(data = Royal_observer_summary, aes(x = counts_per_obs)) +
  geom_histogram(aes_string(fill = "source")) +
  facet_wrap(~ decade_plus, ncol = 1) +
  scale_x_log10() +
  scale_y_continuous(limits = c(0,40)) +
  scale_fill_manual(values = collection_colours, name = "data source") +
  labs(
    y = "number of observers",
    title = "Royal"
    ) +
  theme(
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5),
    legend.position = "none"
  )

histograms <- grid.arrange(
  Royal_histograms, Yos_histograms,
  ncol = 2, 
  widths = c(1, 1.35),
  bottom = textGrob("number of first records", gp = gpar(fontsize = 10))
)

cowplot::save_plot("outputs/observer_histograms.jpg", histograms, base_width = 6, base_height = 9)
```

## New observations 2024.02 to 2024.08
```{r}
iNat_20240901 <- read_csv("data_Yosemite/iNat_Yos_RG_2024.09.01.csv")

iNat_RG_202402 <- Yos_species_paper_2 %>% 
  filter(iNat_RG_END == "1")

unseen_to_RG_summer_24 <- iNat_20240901 %>%
filter(taxon_species_name %in% Yos_species_paper_2$accepted_name) %>%
filter(!taxon_species_name %in% iNat_RG_202402$accepted_name)  %>%
  select(user_login, taxon_species_name, observed_on, latitude, longitude) %>%
  arrange(observed_on) %>%
  distinct(taxon_species_name, .keep_all = TRUE) %>%
  arrange(user_login)

iNat_new_finds_summer_24 <- iNat_20240901 %>%
filter(!taxon_species_name %in% unseen_to_RG_summer_24$taxon_species_name)  %>%
filter(!taxon_species_name %in% iNat_RG_202402$accepted_name) %>%
  select(user_login, taxon_species_name, observed_on, updated_at, latitude, longitude) %>%
  arrange(observed_on) %>%
  distinct(taxon_species_name, .keep_all = TRUE) %>%
  arrange(user_login) %>% View()
```

```{r}
iNat_needsID_20240901 <- read_csv("data_Yosemite/iNat_Yos_needsID_2024.09.01.csv")

iNat_RG_202402 <- Yos_species_paper_2 %>% 
  filter(iNat_RG_END == "1")

unseen_as_RG_thru_summer_24 <- iNat_needsID_20240901 %>%
  arrange(observed_on) %>%
  distinct(taxon_species_name, .keep_all = TRUE) %>%
filter(taxon_species_name %in% Yos_species_paper_2$accepted_name) %>%
filter(!taxon_species_name %in% (Yos_species_paper_2 %>% filter(is.na(recent_voucher)))$accepted_name) %>%
filter(!taxon_species_name %in% iNat_RG_202402$accepted_name)  %>%
filter(!taxon_species_name %in% unseen_to_RG_summer_24$taxon_species_name)  %>%
  select(user_login, taxon_species_name, observed_on, latitude, longitude) %>%
  arrange(user_login) %>%
  left_join(Yos_species_paper_2 %>% rename(taxon_species_name = accepted_name) %>% select(taxon_species_name, establishment_means, family, broader))



```


